---
title: "test-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{test-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview 
This vignette introduces users to each of the four models available to use in the package by fitting all four models to a single simulated data set. 
This is a recreation (with fewer samples) of the analysis shown in Figure 2 in ``Semiparametric Inference of Effective Reproduction Number
Dynamics from Wastewater Pathogen RNA Concentrations." 

# Set up
You will need to have installed julia (at least version 1.8.5) as well as the package [concRt.jl](https://github.com/igoldsteinh/concRt.jl) in order to run this code. 
See the README for instructions on installation. 
This package relies on the JuliaCall package, the following code loads needed libraries and sets up julia so that it can be used within R. 
```{r message = FALSE, warning = FALSE}
# set number of threads (one for each chain)
Sys.setenv(JULIA_NUM_THREADS = "4")
# load packages
library(concRt)
library(JuliaCall)
library(ggplot2)
library(tidyr)
library(tibble)
library(dplyr)
library(cowplot)
library(scales)
library(latex2exp)
library(tidybayes)
library(stringr)
# set up theme by Damon Bayer
my_theme <- list(
  scale_fill_brewer(name = "Credible Interval Width",
                    labels = ~percent(as.numeric(.))),
  guides(fill = guide_legend(reverse = TRUE)),
  theme_minimal_grid(),
  theme())

# you must use julia_setup every time you use this package
# change the address to wherever julia is installed on your computer
# if installed using setup(), you may be able to leave this blank
julia <- julia_setup(JULIA_HOME = "/Users/isaacgoldstein/.juliaup/bin")
# load required julia package 
julia_library("concRt")
# check number of threads, if is not 4, something went wrong. 
# See https://docs.julialang.org/en/v1/manual/multi-threading/#Starting-Julia-with-multiple-threads
julia_eval("Threads.nthreads()")
```

# The EIRR Model
First we load the data and fit the model
```{r}
# transform data into model input
long_dat <- scenario1_genecount_data %>% 
            dplyr::filter(seed == 1) %>% 
            dplyr::select(new_time, log_gene_copies1, log_gene_copies2, log_gene_copies3) %>%             
            tidyr::pivot_longer(-new_time)

data <- long_dat$value
obstimes <- long_dat$new_time
param_change_times <- c(7.0, 14.0, 21.0, 28.0, 35.0, 42.0, 49.0, 
                        56.0, 63.0, 70.0, 77.0, 84.0, 91.0, 98.0, 
                        105.0, 112.0, 119.0, 126.0)
priors_only <- FALSE
n_samples <- 25L
n_chains <- 4L
seed <- 1L

# generate posterior samples  
posterior_samples_eirr <- fit_eirrc(data, 
                               obstimes, 
                               param_change_times, 
                               priors_only, 
                               n_samples, 
                               n_chains, 
                               seed,
                               gamma_sd = 0.2,
                               gamma_mean =log(1/4),
                               nu_sd = 0.2,
                               nu_mean = log(1/7),
                               eta_sd = 0.2,
                               eta_mean = log(1/18),
                               rho_gene_sd = 1.0,
                               rho_gene_mean = 0.0,
                               tau_sd = 1.0,
                               tau_mean = 0.0,
                               I_init_sd = 0.05,
                               I_init_mean = 489.0,
                               R1_init_sd = 0.05,
                               R1_init_mean = 2075.0,
                               E_init_sd = 0.05,
                               E_init_mean = 225.0,
                               lambda_mean = 5.685528,
                               lambda_sd = 2.178852,
                               df_shape = 2.0,
                               df_scale = 10.0,
                               sigma_rt_sd = 0.2,
                               sigma_rt_mean = log(0.1),
                               rt_init_sd = 0.1,
                               rt_init_mean = log(0.88))
```

We then transform the output into quantiles to summarise results
```{r message = FALSE, warning = FALSE}
posterior_output_eirr <- generate_eirrc(posterior_samples_eirr,
                                        data,
                                        obstimes, 
                                        param_change_times,
                                        seed = seed,
                                        gamma_sd = 0.2,
                                        gamma_mean =log(1/4),
                                        nu_sd = 0.2,
                                        nu_mean = log(1/7),
                                        eta_sd = 0.2,
                                        eta_mean = log(1/18),
                                        rho_gene_sd = 1.0,
                                        rho_gene_mean = 0.0,
                                        tau_sd = 1.0,
                                        tau_mean = 0.0,
                                        I_init_sd = 0.05,
                                        I_init_mean = 489.0,
                                        R1_init_sd = 0.05,
                                        R1_init_mean = 2075.0,
                                        E_init_sd = 0.05,
                                        E_init_mean = 225.0,
                                        lambda_mean = 5.685528,
                                        lambda_sd = 2.178852,
                                        df_shape = 2.0,
                                        df_scale = 10.0,
                                        sigma_rt_sd = 0.2,
                                        sigma_rt_mean = log(0.1),
                                        rt_init_sd = 0.1,
                                        rt_init_mean = log(0.88))

eirr_quantiles <- make_timevarying_quantiles(posterior_output_eirr[[2]])

eirr_rt_quantiles <- eirr_quantiles %>% dplyr::filter(name == "rt_t_values")
```

Plot the results. The red dots are the true values, the black line is the posterior median, the blue bars are credible intervals.
```{r fig.width=8, fig.height=6, fig.align='center'}
max_time <- max(scenario1_genecount_data$time)

eirr_plot_quantiles <- eirr_rt_quantiles %>%
  mutate(time = time - 1) %>%
  rename("week" = time) %>%
  right_join(scenario1_fullsimdata, by = "week") %>%
  dplyr::select(week,time, new_time, true_rt, value, .lower, .upper, .width,.point, .interval) %>%
  filter(time <= max_time,
         week >= 0)


eirr_scenario1_rt_plot <- eirr_plot_quantiles %>%
  ggplot(aes(time, value, ymin = .lower, ymax = .upper)) +
  geom_lineribbon() +
  geom_point(aes(time, true_rt), color = "coral1") + 
  scale_y_continuous("Rt", label = comma) +
  scale_x_continuous(name = "Time") +
  ylim(c(0,3.5)) +
  ggtitle(stringr::str_c("EIRR (WW)")) +
  my_theme + 
  theme(legend.position = c(0.6, 0.8),
        legend.background = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  ylab(TeX('$R_{t}$')) 

eirr_scenario1_rt_plot
```

# The SEIRR Model
Again we fit the model to the data. 
```{r}

# transform data into model input
long_dat <- scenario1_genecount_data %>% 
            filter(seed == 1) %>% 
            dplyr::select(new_time, log_gene_copies1, log_gene_copies2, log_gene_copies3) %>%             
            pivot_longer(-new_time)

data <- long_dat$value
obstimes <- long_dat$new_time
param_change_times <- c(7.0, 14.0, 21.0, 28.0, 35.0, 42.0, 49.0, 
                        56.0, 63.0, 70.0, 77.0, 84.0, 91.0, 98.0, 
                        105.0, 112.0, 119.0, 126.0)
priors_only <- FALSE
n_samples <- 25L
n_chains <- 4L
seed <- 1L
extra_ode_precision <- TRUE


posterior_samples_seirr <- fit_seirr(data,
                      obstimes,
                      param_change_times,
                      extra_ode_precision,
                      priors_only,
                      n_samples,
                      n_chains,
                      seed,
                      fit_abs_tol = 1e-9,
                      fit_rel_tol = 1e-6,
                      opt_abs_tol = 1e-11,
                      opt_rel_tol = 1e-8,
                      popsize = 100000L,
                      active_pop = 92271L,
                      gamma_sd = 0.2,
                      gamma_mean =log(1/4),
                      nu_sd = 0.2,
                      nu_mean = log(1/7),
                      eta_sd = 0.2,
                      eta_mean = log(1/18),
                      rho_gene_sd =  1.0,
                      rho_gene_mean = 0.0,
                      tau_sd = 1.0,
                      tau_mean = 0.0,
                      sigma_R0_sd = 0.2,
                      sigma_R0_mean = log(0.1),
                      S_SEIR1_sd = 0.05,
                      S_SEIR1_mean = 3.468354,
                      I_EIR1_sd = 0.05,
                      I_EIR1_mean = -1.548302,
                      R1_ER1_sd = 0.05,
                      R1_ER1_mean = 2.221616,
                      r0_init_sd = 0.2,
                      r0_init_mean = log(0.88),
                      lambda_mean = 5.685528,
                      lambda_sd = 2.178852,
                      df_shape = 2.0,
                      df_scale = 10.0) 

```

Transform the output into quantiles 
```{r message = FALSE, warning = FALSE}
posterior_output_seirr <- generate_seirr(posterior_samples_seirr,
                                        data,
                                        obstimes,
                                        param_change_times,
                                        extra_ode_precision,
                                        seed,
                                        fit_abs_tol = 1e-9,
                                        fit_rel_tol = 1e-6,
                                        popsize = 100000L,
                                        active_pop = 92271L,
                                        gamma_sd = 0.2,
                                        gamma_mean =log(1/4),
                                        nu_sd = 0.2,
                                        nu_mean = log(1/7),
                                        eta_sd = 0.2,
                                        eta_mean = log(1/18),
                                        rho_gene_sd =  1.0,
                                        rho_gene_mean = 0.0,
                                        tau_sd = 1.0,
                                        tau_mean = 0.0,
                                        sigma_R0_sd = 0.2,
                                        sigma_R0_mean = log(0.1),
                                        S_SEIR1_sd = 0.05,
                                        S_SEIR1_mean = 3.468354,
                                        I_EIR1_sd = 0.05,
                                        I_EIR1_mean = -1.548302,
                                        R1_ER1_sd = 0.05,
                                        R1_ER1_mean = 2.221616,
                                        r0_init_sd = 0.2,
                                        r0_init_mean = log(0.88),
                                        lambda_mean = 5.685528,
                                        lambda_sd = 2.178852,
                                        df_shape = 2.0,
                                        df_scale = 10.0)

seirr_quantiles <- make_timevarying_quantiles(posterior_output_seirr[[2]])

seirr_rt_quantiles <- seirr_quantiles %>% filter(name == "rt_t_values")

```

Plot the results. 
The red dots are the true values, the black line is the posterior median, the blue bars are credible intervals.
Unlike all other models, the SEIRR produces estimates of $R_{t}$ on a daily time scale, this is because it solves the ODE at a daily time scale, and so estimates of the fraction of susceptibles on a daily scale are available.
```{r  fig.width=8, fig.height=6, fig.align='center'}
max_time <- max(scenario1_genecount_data$time)

seirr_plot_quantiles <- seirr_rt_quantiles %>%
                      right_join(scenario1_fullsimdata, by = c("time" = "new_time")) %>%
                dplyr::select(time, week, true_rt, value, .lower, .upper, .width,.point, .interval) %>%
                filter(time <= max_time,
                       week >= 0)
                

seirr_scenario1_rt_plot <- seirr_plot_quantiles %>%
  ggplot(aes(time, value, ymin = .lower, ymax = .upper)) +
  geom_lineribbon() +
  geom_point(aes(time, true_rt), color = "coral1") + 
  scale_y_continuous("Rt", label = comma) +
  scale_x_continuous(name = "Time") +
  ylim(c(0,3.5)) +
  ggtitle(str_c("SEIRR (WW)")) +
  my_theme + 
  theme(legend.position = c(0.6, 0.8),
        legend.background = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  ylab(TeX('$R_{t}$')) 

seirr_scenario1_rt_plot
```

# The EIR Model
Now load the simulated case data and fit the model 
```{r}
# transform the data 
long_dat <- scenario1_case_data %>%
    filter(seed == 1)

data_cases <- long_dat$total_cases
obstimes <- long_dat$new_week
param_change_times <- obstimes[1:(length(obstimes) -1)]

priors_only <- FALSE
n_samples <- 25L
n_chains <- 4L
seed <- 1L

# fit the model 
posterior_samples_eir <- fit_eir(data_cases,
                                 obstimes,
                                 param_change_times,
                                 priors_only,
                                 n_samples = n_samples,
                                 n_chains = n_chains,
                                 seed = seed,
                                 gamma_sd = 0.2,
                                 gamma_mean =log(7/4),
                                 nu_sd = 0.2,
                                 nu_mean = log(7/7),
                                 rho_case_sd = 1.0,
                                 rho_case_mean = 0.0,
                                 phi_sd = 0.2,
                                 phi_mean = log(50),
                                 I_init_sd = 0.05,
                                 I_init_mean = 489.0,
                                 E_init_sd = 0.05,
                                 E_init_mean = 225.0,
                                 sigma_rt_sd = 0.2,
                                 sigma_rt_mean = log(0.1),
                                 rt_init_sd = 0.1,
                                 rt_init_mean = log(0.88))
```

Transform the output into quantiles
```{r message = FALSE, warning = FALSE}
posterior_output_eir <- generate_eir(posterior_samples_eir,
                                     data_cases,
                                     obstimes,
                                     param_change_times,
                                     seed = seed,
                                     gamma_sd = 0.2,
                                     gamma_mean =log(1/4),
                                     nu_sd = 0.2,
                                     nu_mean = log(1/7),
                                     rho_case_sd = 1.0,
                                     rho_case_mean = 0.0,
                                     phi_sd = 0.2,
                                     phi_mean = log(50),
                                     I_init_sd = 0.05,
                                     I_init_mean = 489.0,
                                     E_init_sd = 0.05,
                                     E_init_mean = 225.0,
                                     sigma_rt_sd = 0.2,
                                     sigma_rt_mean = log(0.1),
                                     rt_init_sd = 0.1,
                                     rt_init_mean = log(0.88))

eir_quantiles <- make_timevarying_quantiles(posterior_output_eir[[2]])

eir_rt_quantiles <- eir_quantiles %>% filter(name == "rt_t_values")

```

Graph the results
```{r  fig.width=8, fig.height=6, fig.align='center'}
max_time <- max(scenario1_genecount_data$time)

eir_plot_quantiles <- eir_rt_quantiles %>%
  mutate(time = time - 1) %>%
  rename("week" = time) %>%
  right_join(scenario1_fullsimdata, by = "week") %>%
  dplyr::select(week,time, new_time, true_rt, value, .lower, .upper, .width,.point, .interval) %>%
  filter(time <= max_time,
         week >= 0)
                

eir_scenario1_rt_plot <- eir_plot_quantiles %>%
  ggplot(aes(time, value, ymin = .lower, ymax = .upper)) +
  geom_lineribbon() +
  geom_point(aes(time, true_rt), color = "coral1") + 
  scale_y_continuous("Rt", label = comma) +
  scale_x_continuous(name = "Time") +
  ylim(c(0,3.5)) +
  ggtitle(str_c("EIR (Cases)")) +
  my_theme + 
  theme(legend.position = c(0.6, 0.8),
        legend.background = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  ylab(TeX('$R_{t}$')) 

eir_scenario1_rt_plot
```

# The SEIR Model
Again load the case data and fit the model
```{r}
# transform the data 
long_dat <- scenario1_case_data %>%
    filter(seed == 1)

data_cases <- long_dat$total_cases
obstimes <- long_dat$new_week
param_change_times <- obstimes[1:(length(obstimes) -1)]

priors_only <- FALSE
n_samples <- 25L
n_chains <- 4L
seed <- 1L
extra_ode_precision <- TRUE

posterior_samples_seir <- fit_seir(data_cases,
                                   obstimes,
                                   param_change_times,
                                   extra_ode_precision,
                                   priors_only,
                                   n_samples = n_samples,
                                   n_chains = n_chains,
                                   seed = seed,
                                   fit_abs_tol = 1e-9,
                                   fit_rel_tol = 1e-6,
                                   opt_abs_tol = 1e-11,
                                   opt_rel_tol = 1e-8,
                                   popsize = 100000L,
                                   active_pop = 90196L,
                                   gamma_sd = 0.2,
                                   gamma_mean =log(7/4),
                                   nu_sd = 0.2,
                                   nu_mean = log(7/7),
                                   rho_case_sd =  0.4,
                                   rho_case_mean = -1.386294,
                                   phi_sd = 0.2,
                                   phi_mean = log(50),
                                   sigma_R0_sd = 0.2,
                                   sigma_R0_mean = log(0.1),
                                   S_SEI_sd = 0.05,
                                   S_SEI_mean = 4.83091,
                                   I_EI_sd = 0.05,
                                   I_EI_mean = 0.7762621,
                                   r0_init_sd = 0.1,
                                   r0_init_mean = log(0.88))
```

Transform the output into quantiles 
```{r message = FALSE, warning = FALSE}
posterior_output_seir <- generate_seir(posterior_samples_seir,
                                       data_cases,
                                       obstimes,
                                       param_change_times,
                                       extra_ode_precision,
                                       seed = seed,
                                       fit_abs_tol = 1e-9,
                                       fit_rel_tol = 1e-6,
                                       popsize = 100000L,
                                       active_pop = 90196L,
                                       gamma_sd = 0.2,
                                       gamma_mean =log(7/4),
                                       nu_sd = 0.2,
                                       nu_mean = log(7/7),
                                       rho_case_sd =  0.4,
                                       rho_case_mean = -1.386294,
                                       phi_sd = 0.2,
                                       phi_mean = log(50),
                                       sigma_R0_sd = 0.2,
                                       sigma_R0_mean = log(0.1),
                                       S_SEI_sd = 0.05,
                                       S_SEI_mean = 4.83091,
                                       I_EI_sd = 0.05,
                                       I_EI_mean = 0.7762621,
                                       r0_init_sd = 0.1,
                                       r0_init_mean = log(0.88))

seir_quantiles <- make_timevarying_quantiles(posterior_output_seir[[2]])

seir_rt_quantiles <- seir_quantiles %>% filter(name == "rt_t_values")

```

Graph the results 
```{r  fig.width=8, fig.height=6, fig.align='center'}
max_time <- max(scenario1_genecount_data$time)

seir_plot_quantiles <- seir_rt_quantiles %>%
  mutate(time = time - 1) %>%
  rename("week" = time) %>%
  right_join(scenario1_fullsimdata, by = "week") %>%
  dplyr::select(week,time, new_time, true_rt, value, .lower, .upper, .width,.point, .interval) %>%
  filter(time <= max_time,
         week >= 0)
                

seir_scenario1_rt_plot <- seir_plot_quantiles %>%
  ggplot(aes(time, value, ymin = .lower, ymax = .upper)) +
  geom_lineribbon() +
  geom_point(aes(time, true_rt), color = "coral1") + 
  scale_y_continuous("Rt", label = comma) +
  scale_x_continuous(name = "Time") +
  ylim(c(0,3.5)) +
  ggtitle(str_c("SEIR (Cases)")) +
  my_theme + 
  theme(legend.position = c(0.6, 0.8),
        legend.background = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  ylab(TeX('$R_{t}$')) 

seir_scenario1_rt_plot
```

